---
title: "Reproducible Research Project"
date: "3/29/2020"
output: html_document
---

Loading in data from provided csv file
```{r echo=T}
###Loading Data
library(lattice)
x=read.csv("activity.csv",stringsAsF=F)
```
---

Determine the average steps taken per day and plot this distribution as a histogram. This subject usually takes around 10,000 steps a day.
```{r echo=T}
### What is mean total number of steps taken per day?

steps_per_day=tapply(x$steps,x$date,sum,na.rm=T)
hist(steps_per_day, breaks=16, main="Histogram of Steps Per Day",xlab="Total Steps",col="steel blue")

print(paste("Mean: ",mean(steps_per_day)))
print(paste("Median: ",median(steps_per_day)))
```
---

Determine the average steps taken per time interval and plot to visualize effect of daily schedule on activity. The subject clearly has very low activity levels at night while they are presumably sleeping. They have a high spike in activity at around the 800 time interval. This could be a scheduled workout or a commute.
```{r echo=T}
###What is the average daily activity pattern?

mStepsPerInt=tapply(x$steps,x$interval,mean,na.rm=T)#Average steps per Interval
xyplot(mStepsPerInt~as.numeric(names(mStepsPerInt)),type="l",xlab="Time Interval",ylab="Average Steps")
print(paste("Interval with most average steps: ",names(which.max(mStepsPerInt)))) #Interval with highest Mean
```
---

Impute missing values in data using the average value from the corresponding time interval. This notably removes the group of days that had around 0 steps in the previous histogram.

```{r echo=T}
###Imputing missing values

sum(is.na(x$steps)) #number of na values

for(i in seq_along(x$steps)){ #Replace missing values with mean of interval
  if(is.na(x$steps[i])){
    x$steps[i]=mStepsPerInt[which(x[i,3]==names(mStepsPerInt))]
  }
}
sumPerDay2=tapply(x$steps,x$date,sum,na.rm=T)
hist(sumPerDay2,            #plot hist of sum per day again
     main="Histogram of Steps Per Day (w/Imputing)",xlab="Total Steps",col="steel blue",breaks=16)

print(paste("Mean Per Day: ",mean(sumPerDay2))) #average day
print(paste("Median Per Day: ",median(sumPerDay2))) #median day  
```
---


Plot the average steps taken per time interval, separating weekends and weekdays. The subject has a different activity schedule on weekends. Most notably, activity is much lower at the beginning of the day on weekends.
```{r echo=T}
###Are there differences in activity patterns between weekdays and weekends?

x$weekday=weekdays(as.Date(x$date)) #record weekday of each date

x$dayType[x$weekday=="Sunday"|x$weekday=="Saturday"]="weekend"
x$dayType[x$weekday!="Sunday"&x$weekday!="Saturday"]="weekday"
x$dayType=factor(x$dayType,levels=c("weekend","weekday"))

wkAg=aggregate(steps~interval+dayType,x,mean)
with(wkAg,xyplot(steps~interval|dayType,type="l",lwd=1,layout=c(1,2)))
```
